### 2018-08-12 thesis_report/Thesis_Model_Improvements.tex
latest

Having a solid baseline model for our malware detection task together with how laboratory \textit{vs.}\ real-world scenarios change the model outcome, we now take this chapter to present the improvements made in order to obtain a more robust model to detect malware.
We start by describing our first improvement, applying a multi layer model to extract more information regarding a sample.
We then take this enhanced model and increase the number of features to include dynamic content and how it impacted the model's results.

@@

On the previous chapter we ended up with a simple \gls{lr} model $\LR$ that given a set of static imports from a sample, would give the probability of it being malware.
Although that was the end purpose of our work, it does not give any deeper understanding of the sample, other than being or not malware.
With that in mind, we propose a new model that not only provides the probability of a sample being malware, but also the probability of being from a specific malware type.

This new model $\mathcal{E}$ comprises a simple ensemble stacking approach, which instead of using a single \gls{lr} classifier, multiple ones are used, layered into two steps.

The first step (layer $\mathcal{E}_{\mathcal{L}_{0}}$) is composed of $n$ \gls{lr} models, where $n$ is the number of possible classes.
Each model is trained to output the likelihood of sample belonging to one of the $n$ classes, in a \textit{one-vs-all} methodology (\ie\ a sample either belongs to $\mathcal{C}_{n}$ or not), having as input the raw features (\eg\ static imports).

@@

With this new model defined, we now present our approach on selecting the $n$ classes of interest.
This represents another labeling problem, but now instead of having to label between goodware and malware, we have to label the malware as belonging to some subclass.

@@

As mentioned in Chapter \ref{chapter:related_work}, there is no agreed upon naming convention for malware, which translates into different names for the same malware sample.
To minimize this problem, we referenced a tool by Sebastián, M. et al.~\cite{sebastian2016avclass}, AVClass, which was built to normalize a malware sample name into the most likely family, using the names provided by VirusTotal~\cite{tool:virustotal}.

@@

@@

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{lccc}
		\toprule
		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
		\midrule
		Trojan			& 97,054 & 44,329 & 24,176\\
		Other			& 49,443 & 24,126 & 14,750\\
		Worm			& 24,554 & 14,837 & 9,381\\
		Virus			& 21,055 & 12,531 & 6,899\\
		Spyware			& 20,724 & 10,172 & 5,955\\
		Ransom			& 7,761 & 1,924 & 1,160\\
		\bottomrule
	\end{tabular}
	\caption{Samples belonging to each of the 6 malware classes for $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
	\label{tab:dataset_sizes_new}
\end{table}

@@

In the previous chapter, we used static imports as features for our malware detection model.
Although the results are reasonable, the information which can be retrieved from static imports alone is limited.
As an example, if a sample is compressed, encrypted or packed, its behavior cannot be inferred from static imports only.
To overcome these limitations, we resort to more dynamic information provided by Cuckoo.

Given that Cuckoo runs the provided samples inside a virtual machine, it monitors the sample from a dynamic point of view.
This monitoring includes information like the sequence of library calls and their categories and network and file activity.
With this in mind, we now detail what dynamic information we extracted from the reports.

@@

With that in mind, we decide to transform the values to follow a normal distribution.
We do this by using scikit-learn's~\cite{tool:sklearn} \textit{QuantileTransformer} with a normal distribution, which splits the possible values into bins such that the resulting distribution is of type \textit{Gaussian} with a mean of 0.
This way we have a greater number of bins around the mean, allowing for better discrimination, whereas very large values fall into the same bin.

@@

Our second type of dynamic information are the number of library calls, which closely relate to the previous feature.
While \textit{category calls} provide the number of calls for a given category, \textit{library calls} provide the count for each library call, hence being a subset of the previous.

@@

We joined these features $\FF_{dynamic}$ to the labeled samples $\CC_{real}$, obtaining a total of 122,633 labeled samples with the new features $\DD_{dynamic} = \CC_{real} \cap \FF_{dynamic}$.

@@

As with previous results, one notes that from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$ the results are affected at all, when the change between the scenarios is merely in the number of malware samples.
Between $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$ we again note the already seen pattern, the score is lowest when using the most realistic dataset.

@@

Having applied the same cross-validation to our modified model, obtaining interesting results, we now go over to test how our temporal based methodologies are affected.

@@

The results are consistent with was previously seen in Section \ref{section:single_layer_results}, with the added factor that the absolute values are higher, and the tendency to increase is more present as we move forward in time, close to the validation set.

@@

Following the previous evaluation order, we now present the results using our \textit{Present-to-Past} validation methodology.
In Figure \ref{fig:presentpast_improved} we present the \gls{auroc} for each iteration, where higher folds represent older samples. Here we see values of 97\% for $\mathcal{S}_{\strict}$, 98\% for $\mathcal{S}_{\loose}$ and 96\% $\mathcal{S}_{\real}$.

@@

For $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ the score is equally and negatively affected by 4\%.
The jagged curve on both scenarios indicate that slight changes on the \gls{fpr} threshold have significant impact on the \gls{tpr}, this may be caused by the smaller dataset size, which in turn creates uneven folds for malware and goodware.

>>>>

Having a solid baseline model for our malware detection task together with how laboratory \textit{vs.}\ real-world scenarios change the model outcome, we now take this chapter to present the improvements made in order to obtain a more robust model to detect malware.
We start by describing our first improvement, the application of a multi layer model to extract more information regarding a sample.
We then take this enhanced model and increase the number of features to include dynamic content and how it impacted the model's results.

@@

On the previous chapter we ended up with a simple \gls{lr} model $\LR$ that given a set of static imports from a sample, would give the probability of it being malware.
Although that was the goal of our work, it does not give any deeper understanding of the sample, other than being or not malware.
With that in mind, we propose a new model that not only provides the probability of a sample being malware, but also the probability of being from a specific malware type.

This new model $\mathcal{E}$ comprises a simple ensemble stacking approach, that uses multiple \gls{lr} classifiers instead of one, layered into two steps.

The first step (layer $\mathcal{E}_{\mathcal{L}_{0}}$) is composed of $n$ \gls{lr} models, where $n$ is the number of possible classes.
Each model is trained to output the likelihood of a sample belonging to one of the $n$ classes, in a \textit{one-vs-all} methodology (\ie\ a sample either belongs to $\mathcal{C}_{n}$ or not), having as input the raw features (\eg\ static imports).

@@

With this new model defined, we now present our approach on selecting the $n$ classes of interest.
This represents another labeling problem, but now instead of having to label between goodware and malware, we have to label the malware according to some subclass.

@@

As mentioned in Chapter \ref{chapter:related_work}, there is no agreed upon naming convention for malware, which translates into different names for the same malware sample.
To minimize this problem, we used a tool by Sebastián, M. et al.~\cite{sebastian2016avclass}, AVClass, which was built to normalize a malware sample name into the most likely family, using the names provided by VirusTotal~\cite{tool:virustotal}.

@@

Our modifications affected steps \textit{(d)} and \textit{(e)} of the AVClass process, depicted in Figure \ref{fig:avclass}.
In step \textit{(d)}, instead of filtering family name tokens, we filtered based on our malware classes.
Doing this affects step \textit{(e)}, where the alias detection is now for our classes.
The remaining step is identical, but the names are now malware classes instead of family names.

@@

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{lccc}
		\toprule
		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
		\midrule
		Trojan			& 97,054 & 44,329 & 24,176\\
		Other			& 49,443 & 24,126 & 14,750\\
		Worm			& 24,554 & 14,837 & 9,381\\
		Virus			& 21,055 & 12,531 & 6,899\\
		Spyware			& 20,724 & 10,172 & 5,955\\
		Ransom			& 7,761 & 1,924 & 1,160\\
		\midrule
		\midrule
		Malware Total	& 98,582 & 45,306 & 24,658\\
		\bottomrule
	\end{tabular}
	\caption{Samples belonging to each of the 6 malware classes for $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
	\label{tab:dataset_sizes_new}
\end{table}

@@

In the previous chapter, we used static imports as features for our malware detection model.
Although the results were interesting, the information which can be retrieved from static imports alone is limited.
As an example, if a sample is compressed, encrypted or packed, its behavior cannot be inferred from static imports only.
To overcome these limitations, we resort to more dynamic information provided by Cuckoo.

Given that Cuckoo runs the provided samples inside a virtual machine, it monitors the sample from a dynamic point of view.
This monitoring includes information like the \emph{sequence of library calls} and their \emph{categories} and \emph{network} and \emph{file activity}.
With this in mind, we now detail what dynamic information we extracted from the reports.

@@

With that in mind, we decide to transform the values to follow a normal distribution.
We do this by using scikit-learn's~\cite{tool:sklearn} \textit{QuantileTransformer} with a normal distribution, which splits the possible values into bins such that the resulting distribution is of type \textit{Gaussian} with a mean of 0.
This way we have a greater number of bins around the mean, allowing for better discrimination, whereas very large but uncommon values fall into the same bin.

@@

Our second type of dynamic information are the number of library calls, which closely relate to the previous feature.
While \textit{category calls} provide the number of calls for a given category, \textit{library calls} provide the count for each library call.
Each category from the previous feature contains a set of library calls, hence each library call is contained in some category.

@@

Having these three new features $\FF_{dynamic}$, we made sure our labeled samples $\CC_{real}$ all contained the features, hence we joined the datasets into a slightly smaller dataset $\DD_{dynamic} = \CC_{real} \cap \FF_{dynamic}$, containing 122,633 samples.

@@

As with previous results, one notes that from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$ the results are not affected at all, when the change between the scenarios is merely in the number of malware samples.
Between $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$ we again note the already seen pattern, the score is lowest when using the more realistic dataset.

@@

Having applied the same cross-validation to our modified model, and obtaining interesting results, we now go over to test how our temporal based methodologies are affected.

@@

The results are consistent with what was previously seen in Section \ref{section:single_layer_results}, with the added factor that the absolute values are higher, and the tendency to increase is more present as we move forward in time, close to the validation set.

@@

Following the previous evaluation order, we now present the results using our \textit{Present-to-Past} validation methodology.
In Figure \ref{fig:presentpast_improved} we present the \gls{auroc} for each iteration, where higher folds represent the inclusion of older samples. Here we see values of 97\% for $\mathcal{S}_{\strict}$, 98\% for $\mathcal{S}_{\loose}$ and 96\% $\mathcal{S}_{\real}$.

@@

For $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ the score is equally and negatively affected by 4\%.
The jagged curve on both scenarios indicates that slight changes on the \gls{fpr} threshold have significant impact on the \gls{tpr}.
This may be caused by the smaller dataset size, which in turn creates uneven folds for malware and goodware.

### 2018-08-12 thesis_report/Thesis_Practical_Applications.tex
latest

To implement this service, we developed a simple web application that receives files from users and then sends them to a Cuckoo instance running on a server.
We created a custom signature for Cuckoo that takes the analysis results and converts it into features suitable for our model, which is then used to classify the sample.
The output of the model is then appended as a signature the final report.
After analysis our web application extracts basic information regarding the file, as well as the output of the classifier.

>>>>

To implement this service, we developed a simple web application that receives files from users and then sends them to a Cuckoo instance running on a server.
We created a custom signature for Cuckoo that takes the analysis results and converts it into features suitable for our model, which is then used to classify the sample.
The output of the model is then appended as a signature to the final report.
After analysis our web application extracts basic information regarding the file, as well as the output of the classifier.

We implemented our web application using Django~\cite{tool:django}.
The implemented application provides a simple upload form for users to submit their own files.
Upon submission, Django sends the sample to a Cuckoo instance running locally.
While Cuckoo is analyzing the sample, users are redirected to the report page, which shows the current state of analysis.
Upon finishing analysis, the page changes to yield the classifier results.

### 2018-08-12 thesis_report/Thesis_Static_Features.tex
Slight fixes.
latest

Although the cross-validation methodology enables to measure the generalization capabilities of a model, it does not account for temporal ordering of the samples.
Since we want to measure the when training samples predate validation, we now define a couple of temporal based validations, based on the concept of cross-validation:

@@

\paragraph{\textit{Temporal Window}:} This validation, depicted in Figure \ref{fig:sliding_window}, also splits the dataset into $k$ folds, but takes $w$ temporal consistent and contiguous folds (\ie each fold immediately precedes the next one), using the first $w-1$ folds for training (older samples) and the remaining fold (more recent samples) for validation.
By starting with the $w$ first folds of the dataset and sliding one fold on each iteration, we apply a sliding window of size $w$ over the dataset.
For this last validation methodology, we again split the dataset into 10 folds.
The sliding window size $w$, is chosen during the results phase, as its choice depends on previous results.

@@

@@

We define this our single layer model $\LR$ results, as we are directly taking a set of features (static imports in this case) and outputting the probability of a sample being malware.

@@

When directly comparing the average \gls{auroc} for cross-validation and our \textit{Past-to-Present} validation, we note that for both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ the \gls{auroc} remains identical, while for $\mathcal{S}_{\real}$ the score decreases from 0.75\% to 0.68\%.

@@

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{ccccc}
	\hline AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
	\hline Cross-Validation & 0.91 & 0.90 & 0.75 & 90 / 10\\
	\hline Past-to-Present & 0.90 & 0.90 & 0.67 & 10 to 90 / 10\\
	\hline Present-to-Past & 0.90 & 0.90 & 0.69 & 10 to 90 / 10\\
	\hline Sliding-Window & 0.89 & 0.88 & 0.73 & 30 / 10\\
	\hline
\end{tabular}
\caption{Single layer results summary.}
\label{tab:singlelayer_results}
\end{table}

>>>>

Although the cross-validation methodology enables to measure the generalization capabilities of a model, it does not account for temporal ordering of the samples.
Since we want to measure when training samples predate validation, we now define a couple of temporal based validations, based on the concept of cross-validation:

@@

\paragraph{\textit{Temporal Window}:} This validation, depicted in Figure \ref{fig:sliding_window}, also splits the dataset into $k$ folds, but takes $w$ temporal consistent and contiguous folds (\ie\ each fold immediately precedes the next one), using the first $w-1$ folds for training (older samples) and the remaining fold (more recent samples) for validation.
By starting with the $w$ first folds of the dataset and sliding one fold on each iteration, we apply a sliding window of size $w$ over the dataset.
For this last validation methodology, we again split the dataset into 10 folds.
The sliding window size $w$, is chosen during the results phase, as its choice depends on previous results.

@@

We measure the \gls{auroc} during each validation's iteration and use the average measurement to discuss the results.

@@

We define this as our single layer model $\LR$ results, as we are directly taking a set of features (static imports in this case) and outputting the probability of a sample being malware.

@@

When directly comparing the average \gls{auroc} for cross-validation and our \textit{Past-to-Present} validation, we note that for both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ the \gls{auroc} remains identical, while for $\mathcal{S}_{\real}$ the score decreases from 75\% to 67\%.

@@

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.0} % more space between rows
	\centering
	\begin{tabular}{ccccc}
		\toprule
		AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
		\midrule
		Cross-Validation & 0.91 & 0.90 & 0.75 & 90 / 10\\
		Past-to-Present & 0.90 & 0.90 & 0.67 & 10 to 90 / 10\\
		Present-to-Past & 0.90 & 0.90 & 0.69 & 10 to 90 / 10\\
		Sliding-Window & 0.89 & 0.88 & 0.73 & 30 / 10\\
		\bottomrule
\end{tabular}
\caption{Single layer results summary.}
\label{tab:singlelayer_results}
\end{table}
